The low dimensional world of echolocating bats

### 2020-03-15 Paper/main.tex
intermediate

One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects [REF]. However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}. Sonar maps the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears. While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal. In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function [REF]. However, complex environments return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible. Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.

@@

In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently detailed, description of the world. We test this hypothesis in two steps. In a first step, we generate an extensive collection of synthetic echo sequences. We assess the compressibility of these echo signals using \PCA. Next, we filter the echo sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}. The synthetic echo sequences differ greatly in complexity. Nevertheless, we show that they can be compressed considerably. Moreover, the output of the cochlear model can be mapped onto an even lower-dimensional space. In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without loss of information. We do this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.

Together these results suggest that echolocating bats live in a low dimensional perceptual world. The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences. Nevertheless, these features are sufficiently rich for bats to recognize locations, and as we discuss, support flight control.

@@

Each set of points was converted to an impulse response with a sample frequency $f_s$ of 219 kHz. The impulse response was limited to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways) and atmospheric attenuation. The atmospheric attenuation was xx dB/m, corresponding to the attenuation at xx kHz. We did not model any variation in target strength.

\todo[inline]{Implement variations in echo strength. And include frequency specific absorption.}

Finally, the impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz. Again, these parameters correspond to those used in the acoustic data collection described below. Example of the resulting echo sequences are shown in figure \ref{fig:synthetictemplates}.

Artificial echo sequences were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. This model simulates the temporal and spectral resolution of the bat's cochlea. The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression. Finally, the model assumes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells. In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz. The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}. The first 6 ms of data were removed. This was done to ensure the data corresponded with the acoustic data. As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal.

Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence. It also depicts the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model. From these responses, it can be inferred that these are highly correlated. This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig. \ref{fig:processing}xx). From these high cross-channel correlations we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information. An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx. In correspondence to our our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.

@@

%Selected examples of templates are depicted in figure \ref{fig:synthetictemplates}.

%The output of the cochlear model was averaged across frequencies to obtain a time-intensity profile. The templates were downsampled to a sampling rate of 2.8 kHz. This sampling rate corresponds to a temporal integration of 350 $\mu$s, corresponding to what has been reported for bats' auditory systems \citep{Wiegrebe1996,Simmons1989}. This operation reduced the number of samples in a template from 7499 to \prepca\ data points. Finally,

%We used principal component analysis to assess the dimensionality of the artificial templates. We found that almost all of the variance in the templates could be captured by \pca\ components (see figure \ref{fig:synthetictemplates}xx). Some examples of templates reconstructed from the principal components are shown in \ref{fig:synthetictemplates}xx). Panel xx of figure \ref{fig:synthetictemplates}xx) shows that most of the variance not captured by these components results in some additional noise for very low values of the templates. The components are visualized in figure \ref{fig:components}.

In this section, we proposed a simple generative model for naturalistic echoes. We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$) which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bats auditory system, these artificial echoes can be converted to low dimensional descriptions. The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

@@

In this work, we reuse the large library of templates constructed by \citet{Vanderelst2016}. Below, we briefly discuss the construction of the templates, and we refer to our previous work for more details. \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citep{Steckel2013a}. The emitter was a Sensecomp 7000 Series ultrasonic broadband speaker. The ensonification device was mounted on a pan-tilt system allowing to rotate it in azimuth and elevation. The device and pan-tilt system were mounted on a tripod.

@@

For every azimuth and elevation direction at each of the  40 or 50 locations, a biologically plausible acoustic signature of the environment was derived from the echoes, i.e., a template. Below, we briefly outline the steps used to derive templates from the acoustic data. Again, we refer to our previous work for details on this procedure \citep{Vanderelst2016}.

As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. As we concluded from the analyzing the synthetic echoes that the response in different frequency channels is highly correlated, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile. The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}. Therefore, templates were averaged across three neighboring directions. The first 6 ms of data were removed to avoid using emitted signals for analysis. Accordingly, each template consisted of \pca\ points. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

Among the sets of templates, some had very little energy in them, notably those collected for directions without reflectors. For example, aiming the ensonification device upwards often resulted in no echoes returning. These templates could not be classified \citep{Vanderelst2016}. Therefore, we removed them from the current analyses by omitting all templates with very low energy. The resulting number of templates included in the current analyses are given in table \ref{tab:sizes}.

@@

We wished to assess whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table, and discriminability (or recognizability) was assessed based on perfect representations of the templates. The Euclidean distance between two echo signatures was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation require a comprehensive search of memory. Hence, in our previous work, we established that templates could be classified, but this required a significant amount of memory and computational resources.

Here, we used a more biologically plausible memory system: an artificial neural network. These are neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}. However, we acknowledge that current training methods most lickely have no biological counterpart.

@@

We should point that we did not take any steps to avoid so-called overfitting by the neural networks. Overfitting occurs when a neural network \citep{Ghotra2017}, or other mathematical models \citep{Hawkins2004}, extract  complicated relationships between inputs and outputs that do not hold beyond the training examples. Overfitting typically reduces the ability of the model to generalize its predictive power to new data. However, in this case, the model is not required to generalize. Instead, we are only interested whether the neural networks are able to learn and classify the set of training examples.

@@

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{../results/history}
	\caption{Training history of the neural networks. The plots give the value of the loss function as a function of training epoch. Notice the logarithmic x-axis.}
	\label{fig:history}
\end{figure}

Figure \ref{fig:israelperformance} summarizes the performance of the three neural networks trained on the Israel templates. The performance was obtained by assuming that the most active output neuron represented the classification of the network. Inspecting the cumulative distribution of error reveals that the neural network encoding outperformed the perfect memory approach. The number of small errors was larger for the neural network than for the perfect memory. The neural network was able to classify about 70\% of templates with an error of less than 30 degrees in azimuth. In elevation, almost all templates were classified with an error below 15 degrees. Most (70\%) errors in location attribution were smaller than 0.5 meters. The confusion matrices confirm that most classification errors were small.

@@

We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of about \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:templates}). Therefore, both templates derived from synthetic echoes and templates constructed from real echoes map into the same lower dimensional space. This suggest that, in general, the filtering of the auditory periphery greatly reduces the dimensionality of the description of the bat's environment.

@@

Here, we show that a low dimensional signal is sufficient for recognizing environments. This is a critical component of navigation. However, bats do more than navigating. They also use sonar to control flight, for example, avoid obstacles or following guiding structures. The description the bat has access to for flight control is has the same low level character than that used for place recognition. Therefore, our work suggests that this low level description is sufficient for flight control as well. We have shown before that a low level description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. However, in that work, we converted both the echoes at the left and at the right ear to template-like representation. Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One can presume a neural network that converts the \pca\ features to flight control parameters.

@@

We were unable to find numbers on the number of neurons in bat brains. However, the mouse cortex has been estimated to house about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}. The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections. Therefore, in as much as these neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex. Let alone the complete cortex, which has been estimated to contain 75 million neurons \citep{Williams2000}.

Despite their small size, the networks in this paper were able to encode (memorize) the templates as least as well as the the perfect memory approach. The neural networks presented here are tiny compared to the density of cortical matter. However, the networks are also small compared to the data they encode. Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates. The number of floats making up the templates is larger than the number of weights, by a factor of 2-3. Therefore, the networks have a smaller memory footprint than the data they encode. It should be noted that we have not made an attempt to try and to optimize (minimize) the size of the networks.

\begin{table}
	\centering
	\begin{tabular}{lrrrrl}
		\toprule
		Site &       T &     n &      W &  W (nb) &      R \\
		\midrule
		Israel &  179730 &  5991 &  73050 &   72000 &  40.6\% \\
		Royal &  214890 &  7163 &  50745 &   49875 &  23.6\% \\
		\bottomrule
	\end{tabular}
	\label{tab:sizes}
	\caption{Table indicating the size of both template data sets and number of weights in the networks. The template size $T$ is obtained as $\pca \times n$, with $n$ the number of templates in the set. The number of weights $W$ includes all connections between neurons as well as the bias weights for each of the three networks trained. $W (nb)$ omits the bias weights. The ratio $R$ gives the ratio $W/T$.}
\end{table}

@@

%Together these two findings lend additional credibility to our hypothesis that bats could recognize locations using their acoustic signature, operationalized as templates. Before, we demonstrated that templates are highly specific to locations and directions. Here, we show that storing such templates does not require extensive neural hardware. Small neural networks can be used to store many (6000-7000) templates. Moreover, this is feasible because templates can effectively be compressed, which we demonstrated here by employing PCA and neural networks of a limited size.

%It is well know that even small, shallow, neural networks are remarkably good at approximating potentially very complex function, even to the extent that it becomes valid to ask why they so powerful. \citet{Lin2017a} suggested that most datasets we are interested in originate from some simple physical law or some generative model with relatively
%few free parameters. In the case of templates, the physical laws that generate them are quite straightforward. In addition, it can be assumed that the structure of natural environments can be described by generative models. Therefore, the ideas put forward by \citet{Lin2017a} might explain, on an abstract level why templates should not require extensive neural hardware to encode.

@@

\onecolumn
\appendix
\section{Supplementary material}
\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{../results/components}
	\caption{}
	\label{fig:components}
\end{figure}

>>>>

One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016}. However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}. Sonar maps the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears. While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal. In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function [REF]. However, complex environments, and even single plants [REF], return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible. Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.

@@

In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently detailed, description of the world. We test this hypothesis in two steps. In a first step, we generate an extensive collection of synthetic echo sequences. We assess the compressibility of these echo signals using \PCA. Next, we filter the echo sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}. The synthetic echo sequences differ greatly in complexity. Nevertheless, the output of the cochlear model can be mapped onto a low-dimensional space. In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without loss of information. We do this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.

Together these results suggest that echolocating bats live in a low dimensional perceptual world. The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences. Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, support flight control.

@@

Each set of points was converted to an impulse response with a sampling frequency $f_s$ of 219 kHz. The impulse response length was set to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways). We varied the amplitude of the impulses according to a normal distribution with a standard deviation of 10 dB.

The impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz. Again, these parameters correspond to those used in the acoustic data collection described below. Examples of the resulting echo sequences are shown in figures \ref{fig:processing} and \ref{fig:synthetictemplates}.

Artificial echo sequences were filtered using a model of the bat's auditory periphery, as proposed by \citet{Wiegrebe2008}. This model simulates the temporal and spectral resolution of the bat's cochlea. The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression. Finally, the model assumes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells. In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz. The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}.  The ouput of the channels were filtered with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency. This filtering approximated frequency-dependent atmospheric attenuation. The first 6 ms of data were removed. This was done to ensure the data corresponded with the acoustic data. As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal.

Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence. It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model. From these responses, it can be inferred that these are highly correlated. This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig. \ref{fig:processing}xx). From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information. An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx. In correspondence to our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.

@@

In this section, we  used a generative model for naturalistic echoes. We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted to low dimensional descriptions. The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

@@

We reused the large library of echoes gathered in real bat habitats by \citet{Vanderelst2016}. Below, we briefly discuss the construction of templates based on these echoes, and we refer to our previous work for more details. \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citep{Steckel2013a}. The emitter was a Sensecomp 7000 Series ultrasonic broadband speaker. The ensonification device was mounted on a pan-tilt system allowing to rotate it in azimuth and elevation. The device and pan-tilt system was mounted on a tripod.

@@

For every azimuth and elevation direction at each of the  40 or 50 locations, a biologically plausible acoustic signature of the environment was derived from the echoes, i.e., a template. Below, we briefly outline the steps used to derive templates from the acoustic data.

As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated. Therefore, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile, or template. The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}. Therefore, templates were averaged across three neighboring directions. The first 6 ms of data were removed to avoid using emitted signals for analysis. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

Among the sets of templates, some had very little energy in them, notably those collected for directions without reflectors. For example, aiming the ensonification device upwards often resulted in no echoes returning. These templates could not be classified \citep{Vanderelst2016}. Therefore, we removed them from the current analyses by omitting all templates with very low energy. The resulting number of templates included in the present analyses are given in table \ref{tab:sizes}.

@@

We wished to assess whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table. The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation require a comprehensive search of memory. Hence, in our previous work, we established that templates could be classified, but this required a significant amount of memory and computational resources.

Here, we used a more biologically plausible memory system: an artificial neural network. These are a neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}. However, we acknowledge that current training methods most likely have no biological counterpart.

@@

We should point out that we did not take any steps to avoid so-called overfitting by the neural networks. Overfitting occurs when a neural network \citep{Ghotra2017}, or other mathematical models \citep{Hawkins2004}, extract complicated relationships between inputs and outputs that do not hold beyond the training examples. Overfitting typically reduces the ability of the model to generalize its predictive power to new data. However, in this case, the model is not required to generalize. Instead, we are only interested in whether the neural networks can learn and classify the set of training examples.

@@

Figure \ref{fig:israelperformance} summarizes the performance of the three neural networks trained on the Israel templates. The performance was obtained by assuming that the most active output neuron represented the classification of the network. Inspecting the cumulative distribution of error reveals that the neural network encoding outperformed the perfect memory approach. The number of small errors was larger for the neural network than for the perfect memory. The neural network was able to classify about 70\% of templates with an error of fewer than 30 degrees in azimuth. In elevation, almost all templates were classified with an error below 15 degrees. Most (70\%) errors in location attribution were smaller than 0.5 meters. The confusion matrices confirm that most classification errors were small.

@@

\citet{Reijniers2010a} quantified the transmission of auditory information in the cochlea of bats. In that study, the authors adapted the Meddis inner hair cell model \citep{Meddis2006} to assess the amount of information in the acoustic signal preserved by the cochlea. These authors concluded that spectral notches introduced by overlapping echoes are faithfully encoded by auditory nerve
spike patterns. However, quantifying the information transfer rate also found that about half of the information (bits) are lost in the transduction from acoustic energy to spike trains. In the current paper, we expand on this work and quantify the dimension reduction due to cochlear encoding. Moreover, we show that this reduction in complexity still supports an ecologically valid task: sonar-based scene recognition.

We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}). This suggests the filtering of the auditory periphery provides the bat with a greatly reduced acoustic description of its environment. To identify the locus of the compression, we performed a number of additional analyses. In particular, we evaluated whether the low-dimensional description is a property of the echo sequences, the gammatone filters, or the 1 kHz low-pass filter in the \citet{Wiegrebe2008} model.

We generated templates from the synthetic echo sequences using two versions of the \citet{Wiegrebe2008} model; with and without the 1 Khz low-pass filter. Per default, the model includes this filter to model the response characteristics of the inner hair cells \citep{Meddis2006,Reijniers2010a}. By removing this lowpass filter, we can assess to what extent it is responsible for the observed compressibility of the templates. We also generated templates using echo sequences of which the samples were randomly shuffled. This allows inferring whether the low-dimensionality of the templates is a property of the echo signals (before being filtered through the cochlear model).

Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance. From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences increases from xx to xx. Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx.  Inspecting the average spectrum of the templates (fig. \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz. Therefore, it ought to be possible to completely represent templates by sampling them at 2 kHz,  or about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate. This corresponds to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig. \ref{fig:mechanism}). The number of \pcs\ required to represent unshuffled templates is about half of this, i.c., \pca. Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction. The structure of the echo signals also increases the dimensionality of the templates.

\begin{figure}[tb]
	\centering
	\includegraphics[width=1\linewidth]{../results/mechanism}
	\caption{}
	\label{fig:mechanism}
\end{figure}

@@

Here, we show that a low dimensional signal is sufficient for recognizing environments. This is a critical component of navigation. However, bats do more than navigating. They also use sonar to control flight, for example, avoid obstacles or following guiding structures. The description the bat has access to for flight control has the same low-level character than that used for place recognition. Therefore, our work suggests that this low-level description is sufficient for flight control as well. We have shown before that a low-level description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. However, in that work, we converted both the echoes at the left and at the right ear to template-like representation. Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One can presume a neural network that converts the \pca\ features to flight control parameters.

@@

We were unable to find numbers on the number of neurons in the cortex of bat species. However, the mouse cortex has been estimated to contain about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}. The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections. Therefore, in as much as these neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex. Let alone the complete cortex, which has been estimated to contain 75 million neurons \citep{Williams2000}.

Despite their small size, the networks in this paper were able to encode (memorize) the templates as least as well as the perfect memory approach. The neural networks presented here are tiny compared to the density of cortical matter. However, the networks are also small compared to the data they encode. Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates. The number of floats making up the templates is larger than the number of weights by a factor of 2-3. Therefore, the networks have a smaller memory footprint than the data they encode. It should be noted that we have not made an attempt to try and to optimize (minimize) the size of the networks.

\begin{table}
	\centering
\begin{tabular}{lrrrrl}
	\toprule
	Site &       T &     n &      W &  W (nb\} &      R \\
	\midrule
	Israel &  206661 &  5299 &  53322 &   52425 &  25.8\% \\
	Royal &  302991 &  7769 &  51945 &   51075 &  17.1\% \\
	\bottomrule
\end{tabular}
	\label{tab:sizes}
	\caption{Table indicating the size of both template data sets and number of weights in the networks. The template size $T$ is obtained as $\pca \times n$, with $n$ the number of templates in the set. The number of weights $W$ includes all connections between neurons as well as the bias weights for each of the three networks trained. $W (nb)$ omits the bias weights. The ratio $R$ gives the ratio $W/T$.}
\end{table}

@@

@@

\onecolumn
\appendix
\section{Supplementary material}

\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{../results/history}
	\caption{Training history of the neural networks. The plots give the value of the loss function as a function of training epoch. Notice the logarithmic x-axis.}
	\label{fig:history}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{../results/components}
	\caption{}
	\label{fig:components}
\end{figure}
